Re-Allocation of Resources during Releases \titlenote{Copyright note}

### 2013-11-19 paper1/paper1.tex
Script for traversing git DAG is ok now. Updated paper.

\documentclass{acm_proc_article-sp}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{listings}
\usepackage{booktabs}
\begin{document}

\title{Re-Allocation of Resources during Releases \titlenote{Copyright note}}
\numberofauthors{2}
\author{
% 1st. author
\alignauthor
Md Tajmilur Rahman\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{438-932-2288, +1}\\
       \email{mdt\_rahm@encs.concordia.ca}
% 2nd. author
\alignauthor
Peter C. Rigby\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{514-848-2424, +1}\\
       \email{peter.rigby@concordia.ca}
}
\date{20 Oct 2013}
\maketitle
\begin{abstract}
This section will be written at the end.
\end{abstract}
\category{K.6.3}{Software Management}{Software Development, Software Resource Management, Resource Reallocation}
\terms{Experiment, Human Factors, Resource Management, Reallocation}
\keywords{Resource Reallocation, Code Ownership, Defect Density, Software Releases} % NOT required for Proceedings
\section{Introduction}
Software projects are notorious for going over budget and schedule. Rush periods are often get seen before a major release that turn the developers into dinosaurs as Frederick Brooks likens in his benchmark study "The Mythical Man Month" \cite{brooks_mythical}. This "Rush To Release (RTR)" can be prompted either by external forces such as decisions by management to include new features in the release or to release earlier to beat a competitor. Alternatively, the rush may simply be due to inappropriate or unrealistic scheduling. Whatever the reason is it is an obvious. Regardless of the causes, the rush to release stresses developers and often requires developers to work on unusual, high priority or critical areas of the system. In this paper we study how RTR effects project organization and introduces technical debt. We want to propose a method to observe, analyze and summarize the results of revisions found near release. We intend to infer behavior of the developers related to the process around release time. The key research questions that we expect to answer with our methodology are as follows:
\begin{enumerate}

\item What is the release process used by the project?
\item Do developers work on different areas of the system around the time of release?
\item Are there certain areas of the system that receive increased attention (i.e.\ do developers focus on a smaller set of files around releases)?
\end{enumerate}

@@

We observe a the developers working areas to understand the allocation of the resources within an open source software development project to identify that an improper reallocation or inappropriate reorganization causes a disruptive event take place in a software development process. We attempt to identify a project's different release times and calculate the difference between different development periods within a releases to identify which new areas are being worked on and what is the behavior of the developers between normal development period and the period when RTR is prompted before a releases. The historical data from git that we are working on for this purpose will help us to extract a lot of information like calculating the developers' working areas and time-frame of each release as well as different periods in a stable release period. This information will help us to identify the criteria of the resources, their roles and behaviors, code ownership and defect density of the domains they are working in.

@@

@@

\subsection{Data Extraction}
We went for the VCS of a target project and either mirror the repository or download every revision and commit log history data. From DVCSs such as Git we extract the revisions and release information. We wrote Perl scripts to extract date that was further processed to obtain details. Manual inspection was used to resolve problems and things like that in cases where all automated techniques failed.
We then put them into a database. We have used PSQL database to create tables in, to store our extracted data. These extracted data will be analyzed by us later on. Per each revision the information extracted includes the commit id, tree id, author of the revision, date of revision, the name of the revised file, parent and child info for the revision and the detail log information. Once extraction is completed we are ready to partition the version numbers (Section 4.2) and duration  of each release (Section 4.3).

@@

@@

We are getting the release dates with every single commit record. So we can understand which commit belongs to which version of release, is this a major release or minor or micro. Another information we have captured that is \textbf{rc} which means that the particular commit was a release candidate.

@@

\subsubsection{DA in a Release Period}
We select all the release information from the table where we stored the releases with their corresponding time-span. After that we select author name and other relevant information from the joining of ``git\_revision'' and ``git\_commits'' tables for every release. We insert these information into another table so that we can use the data for our further analysis.

The following pseudo code shows us the way of achieving the desired result in this section:

\begin{algorithm}
\caption{Calculate Developers Area}
\begin{algorithmic}
\REQUIRE $ $\\
$commit\_date \geq  release\_start\_date$\\
$commit\_date < release\_end\_date$
\ENSURE
\STATE $get all releases$

\WHILE{$prev\_release\_date \leftarrow release\_start\_date\ and\ release \leftarrow release\_version$}
\IF{$release$}
	\IF{$commit\_date \geq  release\_start\_date$ and $commit\_date < release\_end\_date$}
		\STATE $select author, path, commits, churn$
		\STATE $insert data into table dev\_area\_rel$
	\ENDIF
\ENDIF
\ENDWHILE
\end{algorithmic}
\end{algorithm}

A relatively straightforward discipline of Linux Kernel Development team is followed with regard to the merging of patches for each release \cite{linux_kernel}.  At the beginning of each development cycle, the "merge window" is said to be opened.  At that time, code which is deemed to be sufficiently stable (and which is accepted by the development community) is merged into the mainline kernel. The bulk of changes for a new development cycle (and all of the major changes) will be merged during this time, at a rate approaching 1,000 changes ("patches," or "changesets") per day. Strategically linux starts a new kernel release with the merging and fixing to get their development branch ready to start development for the next release. So there are two main segments in a release period, one is merge window or merge period (MP) and another is development window what we are going to call Release Development Period (RDP).

>>>>

\documentclass{acm_proc_article-sp}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{listings}
\usepackage{booktabs}
\usepackage{graphicx}
\begin{document}

\title{Re-Allocation of Resources during Releases \titlenote{Copyright note}}
\numberofauthors{2}
\author{
% 1st. author
\alignauthor
Md Tajmilur Rahman\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{438-932-2288, +1}\\
       \email{mdt\_rahm@encs.concordia.ca}
% 2nd. author
\alignauthor
Peter C. Rigby\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{514-848-2424, +1}\\
       \email{peter.rigby@concordia.ca}
}
\date{20 Oct 2013}
\maketitle
\begin{abstract}
This section will be written at the end.
\end{abstract}
\category{K.6.3}{Software Management}{Software Development, Software Resource Management, Resource Reallocation}
\terms{Experiment, Human Factors, Resource Management, Reallocation}
\keywords{Resource Reallocation, Code Ownership, Defect Density, Software Releases} % NOT required for Proceedings
\section{Introduction}
Software projects are notorious for going over budget and schedule. Rush periods are often get seen before a major release that turn the developers into dinosaurs as Frederick Brooks likens in his benchmark study "The Mythical Man Month" \cite{brooks_mythical}. This "Rush To Release (RTR)" can be prompted either by external forces such as decisions by management to include new features in the release or to release earlier to beat a competitor. Alternatively, the rush may simply be due to inappropriate or unrealistic scheduling. Whatever the reason is it is an obvious. Regardless of the causes, the rush to release stresses developers and often requires developers to work on unusual, high priority or critical areas of the system. In this paper we study how RTR effects project organization and introduces technical debt. We want to propose a method to observe, analyze and summarize the results of revisions found near release. We intend to infer behavior of the developers related to the process around release time. The key research questions that we expect to answer with our methodology are described in section 1.1.

@@

We observe the developers' working areas to understand the allocation of the resources within an open source software development project to identify that an improper reallocation or inappropriate reorganization causes a disruptive event take place in a software development process. We attempt to identify a project's different release times and calculate the difference between different development periods within a releases to identify which new areas are being worked on and what is the behavior of the developers between normal development period and the period when RTR is prompted before a releases. The historical data from git that we are working on for this purpose will help us to extract a lot of information like calculating the developers' working areas and time-frame of each release as well as different periods in a stable release period. This information will help us to identify the criteria of the resources, their roles and behaviors, code ownership and defect density of the domains they are working in.

@@

\subsection{Research Questions}
We have set our research questions focusing the key properties of software development process, allocation of developers throughout the process, especially for open source software development. First couple of research questions are to understand the basic structure and strategies of the development and release and merging process practiced by Linux. Rest of the questions are to understand the distribution of developers working areas and the behavior of developers in different segments of a release period.

\renewcommand{\labelenumi}{Q\theenumi:}
\begin{enumerate}
\item What is the release process used by the project? \newline
We are going to give a qualitative description of Linux Kernel development in brief to give anser to this question.
\item How many developers are allocated troughout different segments of a release period? \newline
Answer to this question will give us a big picture of Linux Kernel development community as well as the contrubution of developers in different periods within a stable release period.
\item Do developers work on different areas of the system around the time of release? \newline
We want to see if developers are looking at different types of files in the code-base around the time of release which is being considered as the merge period or merge window\cite{linux_kernel}. We are expecting this would be the RTR are in a release time where technical debt may be introduced.
\item How do developers own code around t he system?
Many developers are working on many files. We will find out which files has been churned mostly by which developer so that we can understand how developers are distributed among the system.
\item Are there certain areas of the system that receive increased attention (i.e.\ do developers focus on a smaller set of files around releases)? \newline
To answer this question we will look into the frequency of file churns made into files. Difference or distance of the files between normal development period and the RTR period or merge period of a release.
\end{enumerate}

@@

\subsection{Data Extraction}
We went for the VCS of a target project and either mirror the repository or download every revision and commit log history data. From DVCSs such as Git we extract the revisions and release information. We wrote Perl scripts to extract date that was further processed to obtain details. Manual inspection was used to resolve problems and things like that in cases where all automated techniques failed.
We then put them into a database. We have used PSQL database to create tables in, to store our extracted data. These extracted data will be analyzed by us later on. Per each revision the information extracted includes the commit id, tree id, author of the revision, date of revision, the name of the revised file, parent and child info for the revision and the detail log information. Once extraction is completed we are ready to partition the version numbers (Section 4.3) and duration  of each release (Section 4.4).

@@

\subsection{Finding Release Commits}
As out first step we had to sort out all the commits together those are belonging to any of the releases that we have. We have all 400441 commits in our hand but though we can't plot them into a single time domain to say that these commits are for release 2.6.12 and these are for 2.6.13 and these are for 2.6.14 and so on because commits go on parallelly in git if there are multiple branches. If there are multiple branches, surely there are multiple merges and the commits prior to a crisscross merge are all ancestors of the heads of all subsequent branches \cite{bird_git}. Figure 1 shows how a git DAG looks like having so many crisscross merges in a busy branching snap-shot of the git DAG of Linux Kernel development.
\begin{figure}
\begin{center}
\includegraphics[height=3.5in,width=2.5in]{gitdag.png}
\caption{\small \sl Branching in git DAG for linux kernel development.}
\end{center}
\end{figure}
So it is possible to make a commit months ago from today while several releases may have already been made in the mean time but that commit can belong to any future release. Commits that are being made to a branch may get merged with other branch or the head branch during the release period. This is why we had to track all the way down the trees for each of the 340 releases in the git DAG to find out which commit contributes to which release. We could do this using Git DAG Traversing (GDT) algorithm as shown in Algorithm 1. After applying GDT we collected 381152 out of 400441 commits those are well organized in the git DAG for Linux Kernel development. As we mentioned earlier that we have last couple of release candidates and the release date for ``linuxv2.6.12'' and first couple of release candidates for ``linuxv3.12'' and we are choping out them because we don't have complete data for version 2.6.12-rc as well as 3.12-rc releases, 19289 commits belonging to those releases. So finally we have 381152 commits for 340 releases in our hand to progress our work on.

\begin{algorithm}
\caption{Traverse Git DAG to find Commits within Releases}
\begin{algorithmic}[1]
\REQUIRE
\STATE commit id for all releases
\ENSURE
\FOR{all commits}
\STATE start from new commit, release R
\STATE store commit in ``git\_commit\_release''
\IF{has parent commit}
	\FOR{each parent commit}
		\IF{if parent commit has a release tag}
			\STATE repeat step 1
		\ELSIF{parent commit already stored}
			\STATE compare release versions
			\IF{release version not smaller than R}
				\STATE update existing\_release $\gets$ R
			\ENDIF
		\ELSE
			\STATE store parent commit in ``git\_commit\_release''
			\STATE commit id $\gets$ parent commit id
			\STATE repeat step 3
		\ENDIF
	\ENDFOR
\ELSIF{no parent commit}
	\STATE repeat step 1
\ENDIF
\ENDFOR
\end{algorithmic}
\end{algorithm}

@@

We are getting the release dates with every single commit record. So we can understand which commit belongs to which version of release; is this a major release or minor or micro. Another information we have captured that is \textbf{rc} which means that the particular commit was a release candidate.

@@

\subsubsection{DA in a Release Period}
We select all the release information from the table where we stored the releases with their corresponding time-span. After that we select author name and other relevant information from the joining of ``git\_revision'' and ``git\_commits'' tables for every release. We insert these information into another table so that we can use the data for our further analysis. A relatively straightforward discipline of Linux Kernel Development team is followed with regard to the merging of patches for each release \cite{linux_kernel}.  At the beginning of each development cycle, the "merge window" is said to be opened.  At that time, code which is deemed to be sufficiently stable (and which is accepted by the development community) is merged into the mainline kernel. The bulk of changes for a new development cycle (and all of the major changes) will be merged during this time, at a rate approaching 1,000 changes ("patches," or "changesets") per day. Strategically linux starts a new kernel release with the merging and fixing to get their development branch ready to start development for the next release. So there are two main segments in a release period, one is merge window or merge period (MP) and another is development window what we are going to call Release Development Period (RDP).
